---
title: "Interactive Rmarkdown Example"
author: "Jared Parrish PhD"
date: "12/20/2021"
runtime: shiny
output: 
  html_document:  
    toc: true
    toc_float: true  
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(lubridate)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(shiny)

options(scipen=999) # remove scientific notation

#Connect to API, Read in data, and clean a bit

url <- "https://data.alaska.gov/resource/7vre-awpz.json"
t_dat <- jsonlite::fromJSON(url)

t_dat <- t_dat %>% mutate_at(c("year","birth_rate","lower_confidence_limit","upper_confidence_limit"), as.numeric)
```

# Overview

Often I find myself wanting to do more with Rmarkdown to enhance the experience for the user. This is mostly when I have need something like a static Rmarkdown HTML document but some of the functionality of a shiny application. One way to achieve this is to add the runtime option in the YAML and specify "shiny".   

The benefit is that you can take the look and feel of a Rmarkdown but add in widgets. The drawback is that these are no longer static and actually shiny apps so have to be hosted.  

For more details see <https://shiny.rstudio.com/articles/interactive-docs.html>.

For this simple example we're going to use data available on the Alaska Open data Portal located at: <https://data.alaska.gov/>  Specifically, we will connect to the "Teen Birth Rates Age Group 15-19 in AK by Borough 2003 through 2018" API in the "Health" portal.  

These data were originally obtained from the CDC and are now hosted on the Open data Portal by Russ Schmid.  

---
# Statewide  

Based on the data available I fist calculated the annual average teen birth rate. It should be noted that this simply averages the provided birth rates for each county in Alaska by year and does not calculate the actual crude Alaska teen birth rate.


```{r echo=FALSE, message=FALSE, warning=FALSE}
st_av <- t_dat %>% group_by(year) %>% summarise(
                   avg.birth.rate = mean(birth_rate),
                   avg.birth.rate.lower = mean(lower_confidence_limit),
                   avg.birth.rate.upper = mean(upper_confidence_limit)
                   )

```

## Static plot and table

First, I might present the trend in a static figure and a table, and add some text about the image and maybe even a statistical test.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height= 4}

st_pt <- ggplot(data = st_av, aes(x = year, y = avg.birth.rate)) +
         geom_ribbon(aes(ymin = avg.birth.rate.lower,ymax = avg.birth.rate.upper),
                     fill = "blue",alpha = 0.25) +
         geom_line(size = 1, color = "#0a306a") +
         geom_point(aes(x = year, y = avg.birth.rate,text = paste("Year:",year,"<br>Teen Birth Rate:",avg.birth.rate,"<br>LowerCI:",avg.birth.rate.lower,"<br>UpperCI:",avg.birth.rate.upper)),size = 2, color = "#0a306a") +
         scale_x_continuous(breaks = c(2003:2018)) +
         ggtitle("Teen Birth Rate (averaged by county), Alaska, 2003-2018") +
         ylab("Average teen birth rate") +
         theme_minimal()
st_pt
htmlTable::htmlTable(as.data.frame(st_av))

```

## Minimal interaction
I might make the graphic a little more "fancy" by using plotly that adds a tooltip 
and some limited interactions and an interactive table.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 9, fig.height= 4}
ggplotly(st_pt, tooltip = c("text"))  
datatable(data = as.data.frame(st_av),extensions = c("Buttons"))
```  

# Region 
Now I want to present these data by region. With a traditional Rmarkdown I'd probably present for each region in a "long" document with a nested table of contents. 

## Region Static 
A quick example of how this could look using three of the 25 counties.  

### Anchorage  
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 9, fig.height= 4}
st_ank <- t_dat %>% filter(county == "Anchorage") %>% 
                    group_by(year) %>%
                    summarise(
                    avg.birth.rate = mean(birth_rate),
                    avg.birth.rate.lower = mean(lower_confidence_limit),
                    avg.birth.rate.upper = mean(upper_confidence_limit)
                    )

st_pt_ank <- ggplot(data = st_ank, aes(x = year, y = avg.birth.rate)) +
         geom_ribbon(aes(ymin = avg.birth.rate.lower,ymax = avg.birth.rate.upper),
                     fill = "blue",alpha = 0.25) +
         geom_line(size = 1, color = "#0a306a") +
         geom_point(aes(x = year, y = avg.birth.rate,text = paste("Year:",year,"<br>Teen Birth Rate:",avg.birth.rate,"<br>LowerCI:",avg.birth.rate.lower,"<br>UpperCI:",avg.birth.rate.upper)),size = 2, color = "#0a306a") +
         scale_x_continuous(breaks = c(2003:2018)) +
         ggtitle("Teen Birth Rate, Anchorage Alaska, 2003-2018") +
         ylab("Teen birth rate") +
         theme_minimal()
ggplotly(st_pt_ank, tooltip = c("text"))  
datatable(data = as.data.frame(st_ank),extensions = c("Buttons"))
```

### North Slope
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 9, fig.height= 4}
st_ns <- t_dat %>% filter(county == "North Slope") %>% 
                    group_by(year) %>%
                    summarise(
                    avg.birth.rate = mean(birth_rate),
                    avg.birth.rate.lower = mean(lower_confidence_limit),
                    avg.birth.rate.upper = mean(upper_confidence_limit)
                    )

st_pt_ns <- ggplot(data = st_ns, aes(x = year, y = avg.birth.rate)) +
         geom_ribbon(aes(ymin = avg.birth.rate.lower,ymax = avg.birth.rate.upper),
                     fill = "blue",alpha = 0.25) +
         geom_line(size = 1, color = "#0a306a") +
         geom_point(aes(x = year, y = avg.birth.rate,text = paste("Year:",year,"<br>Teen Birth Rate:",avg.birth.rate,"<br>LowerCI:",avg.birth.rate.lower,"<br>UpperCI:",avg.birth.rate.upper)),size = 2, color = "#0a306a") +
         scale_x_continuous(breaks = c(2003:2018)) +
         ggtitle("Teen Birth Rate, North Slope Alaska, 2003-2018") +
         ylab("Teen birth rate") +
         theme_minimal()
ggplotly(st_pt_ns, tooltip = c("text"))  
datatable(data = as.data.frame(st_ns),extensions = c("Buttons"))
```

### Juneau
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 9, fig.height= 4}
st_jn <- t_dat %>% filter(county == "Juneau") %>% 
                    group_by(year) %>%
                    summarise(
                    avg.birth.rate = mean(birth_rate),
                    avg.birth.rate.lower = mean(lower_confidence_limit),
                    avg.birth.rate.upper = mean(upper_confidence_limit)
                    )

st_pt_jn <- ggplot(data = st_jn, aes(x = year, y = avg.birth.rate)) +
         geom_ribbon(aes(ymin = avg.birth.rate.lower,ymax = avg.birth.rate.upper),
                     fill = "blue",alpha = 0.25) +
         geom_line(size = 1, color = "#0a306a") +
         geom_point(aes(x = year, y = avg.birth.rate,text = paste("Year:",year,"<br>Teen Birth Rate:",avg.birth.rate,"<br>LowerCI:",avg.birth.rate.lower,"<br>UpperCI:",avg.birth.rate.upper)),size = 2, color = "#0a306a") +
         scale_x_continuous(breaks = c(2003:2018)) +
         ggtitle("Teen Birth Rate, Juneau Alaska, 2003-2018") +
         ylab("Teen birth rate") +
         theme_minimal()
ggplotly(st_pt_jn, tooltip = c("text"))  
datatable(data = as.data.frame(st_jn),extensions = c("Buttons"))
```


## Region interaction
A long file can be a bit of a headache. We can create a few widgets to make this 
a little better experience.  

```{r echo = FALSE}
selectInput("county", label = "Select County:",
              choices = c(levels(as.factor(t_dat$county))), selected = "Anchorage")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 9, fig.height= 4}
epdta <- reactive ({
  varF <- as.character(input$county)
  dat1 <- t_dat %>% filter(county %in% c(varF)) %>% 
                    group_by(year) %>%
                    summarise(
                    avg.birth.rate = mean(birth_rate),
                    avg.birth.rate.lower = mean(lower_confidence_limit),
                    avg.birth.rate.upper = mean(upper_confidence_limit)
                    )
  list(varF = varF, dat1 = dat1)
})  

renderPlotly({
   tdat1 <- epdta()$dat1
    varF <- epdta()$varF
plt1 <- ggplot(data = tdat1, aes(x = year, y = avg.birth.rate)) +
         geom_ribbon(aes(ymin = avg.birth.rate.lower,ymax = avg.birth.rate.upper),
                     fill = "blue",alpha = 0.25) +
         geom_line(size = 1, color = "#0a306a") +
         geom_point(aes(x = year, y = avg.birth.rate,text = paste("Year:",year,"<br>Teen Birth Rate:",avg.birth.rate,"<br>LowerCI:",avg.birth.rate.lower,"<br>UpperCI:",avg.birth.rate.upper)),size = 2, color = "#0a306a") +
         scale_x_continuous(breaks = c(2003:2018)) +
         ggtitle(paste0("Teen Birth Rate,"," ",varF," ", "Alaska, 2003-2018")) +
         ylab("Teen birth rate") +
         theme_minimal()

ggplotly(plt1, tooltip = c("text"))  

})

DT::renderDataTable({
  tdat1 <- epdta()$dat1
  datatable(data = as.data.frame(tdat1),extensions = c("Buttons"))
})
```

Once using the shiny runtime environment the rmarkdown now must be hosted or shared
with the code to someone with the same access to the data and R as it is now a rshiny application.  

Happy Holidays!
