---
title: "R Markdown Tips & Tricks"
author: "Jared Parrish, PhD"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Horsepower and MPG {#custom-id1}  

```{r}
#| include = FALSE
library(dplyr)
library(ggplot2)
```

Sometimes you might want to include the code chunks, but they may take up a lot of space and 
make the document long. You can include "code folding" into the YMAL

```{r}
mtcars %>% 
  ggplot(aes(x = hp, y = mpg)) +
  geom_point()
```


## Additional folding options  
This first option is nice when you have a lot of code and output to the console you want
to include but want to hide by default.
```{r}
#| collapse = TRUE
mtcars %>% 
  glimpse()
```
This option prints the output of glimpse() but hides the code.
```{r}
#| collapse = FALSE
mtcars %>% 
  glimpse()

mtcars %>%
  ggplot(aes(x = cyl, y = hp)) +
  geom_point()
```

## Using Tabsets {.tabset}    
Sometimes you may want to include tabs opposed to a long linear document. You can do this 
by simply specifying a new section heading with "{.tabset}" included. The nested sections will then be tabs in the document.  

### Plot 
```{r}
mtcars %>% 
  ggplot(aes(x = hp, y = mpg)) +
  geom_point()
```

### Table

```{r}
#| collapse = FALSE
mtcars %>% 
  glimpse()
```

## Multiple column page  

Often it is useful to display information in multiple columns on a page. This can be done with adding ":::columns" to the document.

::: columns
::: column
When you specify the ":::column" this will be what is on the left side.  

```{r}
mtcars %>%
  ggplot(aes(x = hp, y = mpg)) +
  geom_point()
```

:::

::: column
This will be what is on the right side of the page.  
```{r}
#| collapse = TRUE
mtcars %>%
  glimpse()
```
:::
:::

You can close out and return to normal full page spanning by ending the "columns" with a ":::". Displaying a table or text next to a plot is a common use of multiple columns.  

## Internal links  


* Go to [Horsepower and MPG](#custom-id1)


